A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex
more results

This section shows performance results for the various algorithms described in section \ref{sec:algs}.

@@

@@

@@

For this experiment we will only use a portion of the train data to speed up the process.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c | c |}
    \hline
    \textbf{Parameter} & \textbf{Portion} & \textbf{Portion error rate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 3 & 5 & & & & \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 5 & 5 & & & & \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 15 & 5 & & & & \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Genetic Search.}
  \label{tbl:results_bayesian_networks_gs}
  \end{center}
\end{table}

@@

The best configuration for hill climbing (Table \ref{tbl:results_bayesian_networks_hc})
is when the number of parents is exactly 1, still, the error rate is somewhat unsatisfactory when
comparing to other results.

@@

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    HillClimber -P 1 & 19.22 & 22.93 & 0m29.512s \\ \hline
    HillClimber -P 2 & 15.15 & 26.34 & 0m30.340s \\ \hline
    HillClimber -P 3 & 14.51 & 28.11 & 0m29.963s \\ \hline
    HillClimber -P 4 & 14.43 & 27.92 & 0m29.945s \\ \hline
    HillClimber -P 10 & 14.43 & 27.92 & 0m30.638s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Hill Climbing.}
  \label{tbl:results_bayesian_networks_hc}
  \end{center}
\end{table}

@@

We tried various temperatures values and came up with the Table \ref{tbl:results_bayesian_networks_sa}.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    SimulatedAnnealing -U 75 -D 0.4 -A 20 & 14.32 & 26.80 & 0m29.841s \\ \hline
    SimulatedAnnealing -U 55 -D 0.4 -A 20 & 14.40 & 21.45 & 0m29.77s \\ \hline
    SimulatedAnnealing -U 52 -D 0.4 -A 20 & 14.49 & 21.09 & 0m29.407s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 20 & 15.2 & 21.35 & 0m30.884s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 40 & 15.2 & 21.35 & 0m30.884s \\ \hline
    SimulatedAnnealing -U 25 -D 0.4 -A 20 & 18.25 & 22.31 & 0m29.602s \\ \hline
    SimulatedAnnealing -U 10 -D 0.4 -A 20 & 19.82 & 22.96 & 0m29.131s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Simulated Annealing.}
  \label{tbl:results_bayesian_networks_sa}
  \end{center}
\end{table}

The diminish the error rates more research will be required to find the best
tuple of configuration values. Anyway, with just a few guesses we came up
with $\approx21$\% error rate, which is quite good when comparing to other algorithms.

@@

>>>>

This section shows performance results for the various algorithms described in section \ref{sec:algs}.
For every algorithm we will considerate the following aspects: train error rate, test error rate and time.
Each error rate is displayed as a percentage.

@@

With the discretized data set this algorithm performs quite badly
against other algorithms with the same data, like C4.5, with $\approx$25\% error rate.

@@

Again, we will use two train sets: a discretized set and a raw set with numeric values.
The bayesian algorithm in Weka, when fed with numeric values,
it auto discretizes the values, as the nature of bayesian algorithms demands.

@@

For this experiment we will only use 1\% of the train data to speed up the process.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c | c |}
    \hline
    \textbf{Parameter} & \textbf{Data} & \textbf{Portion error rate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 3 & discretized & 13.50 & 18.72 & 22.69 & 0m25.115s \\ \hline
    GeneticSearch -C -M -L 3 -A 3 -U 5 & raw & 13.80 & 18.23 & 21.26 & 0m3.601s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Genetic Search.}
  \label{tbl:results_bayesian_networks_gs}
  \end{center}
\end{table}

@@

The best configuration for hill climbing (tables \ref{tbl:results_bayesian_networks_hc_discr}
and \ref{tbl:results_bayesian_networks_hc_raw})
is when the number of parents is exactly 1.

@@

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    HillClimber -P 1 & 19.22 & 22.93 & 0m29.512s \\ \hline
    HillClimber -P 2 & 15.15 & 26.34 & 0m30.340s \\ \hline
    HillClimber -P 3 & 14.51 & 28.11 & 0m29.963s \\ \hline
    HillClimber -P 4 & 14.43 & 27.92 & 0m29.945s \\ \hline
    HillClimber -P 10 & 14.43 & 27.92 & 0m30.638s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Hill Climbing with discretized data.}
  \label{tbl:results_bayesian_networks_hc_discr}
  \end{center}
\end{table}

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    HillClimber -P 1 & 16.61 & 20.03 & 0m6.873s \\ \hline
    HillClimber -P 2 & 13.79 & 23.20 & 0m7.023s \\ \hline
    HillClimber -P 3 & 12.86 & 25.14 & 0m7.166s \\ \hline
    HillClimber -P 4 & 12.54 & 25.60 & 0m8.326s \\ \hline
    HillClimber -P 10 & 12.54 & 25.60 & 0m9.236s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Hill Climbing with raw data.}
  \label{tbl:results_bayesian_networks_hc_raw}
  \end{center}
\end{table}

@@

We tried various temperatures values and came up with the tables \ref{tbl:results_bayesian_networks_sa_discr} and
\ref{tbl:results_bayesian_networks_sa_raw}.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    SimulatedAnnealing -U 75 -D 0.4 -A 20 & 14.32 & 26.80 & 0m29.841s \\ \hline
    SimulatedAnnealing -U 55 -D 0.4 -A 20 & 14.40 & 21.45 & 0m29.77s \\ \hline
    SimulatedAnnealing -U 52 -D 0.4 -A 20 & 14.49 & 21.09 & 0m29.407s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 20 & 15.2 & 21.35 & 0m30.884s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 40 & 15.2 & 21.35 & 0m30.884s \\ \hline
    SimulatedAnnealing -U 25 -D 0.4 -A 20 & 18.25 & 22.31 & 0m29.602s \\ \hline
    SimulatedAnnealing -U 10 -D 0.4 -A 20 & 19.82 & 22.96 & 0m29.131s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Simulated Annealing with discretized data.}
  \label{tbl:results_bayesian_networks_sa_discr}
  \end{center}
\end{table}

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    SimulatedAnnealing -U 75 -D 0.4 -A 20 & 12.90 & 22.74 & 0m7.270s \\ \hline
    SimulatedAnnealing -U 55 -D 0.4 -A 20 & 13.53 & 19.15 & 0m7.223s \\ \hline
    SimulatedAnnealing -U 52 -D 0.4 -A 20 & 13.55 & 19.15 & 0m7.712s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 20 & 13.55 & 19.15 & 0m7.164s \\ \hline
    SimulatedAnnealing -U 50 -D 0.4 -A 40 & 13.55 & 19.15 & 0m7.252s \\ \hline
    SimulatedAnnealing -U 25 -D 0.4 -A 20 & 15.85 & 19.90 & 0m7.239s \\ \hline
    SimulatedAnnealing -U 10 -D 0.4 -A 20 & 16.60 & 19.91 & 0m7.168s \\ \hline
  \end{tabular}
  \caption{Results for Bayesian Networks / Simulated Annealing with raw data.}
  \label{tbl:results_bayesian_networks_sa_raw}
  \end{center}
\end{table}

The diminish the error rates more research will be required to find the best
tuple of configuration values. Anyway, with just a few guesses we came up
with $\approx19.15$\% error rate, which is quite good when comparing to other algorithms.
The Bayesian network can be seen in Listing \ref{bayes_net_example}.

@@

\clearpage
\section{Bayesian Networks}

\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Bayesian Network built using a Simulated Annealing search algorithm},label=bayes_net_example]
Bayes Network Classifier
Using ADTree
#attributes=13 #classindex=12
Network structure (nodes followed by parents)
age(9): plus_50 workclass 
workclass(9): plus_50 
education(16): age plus_50 
marital_status(7): plus_50 
occupation(15): plus_50 
relationship(6): plus_50 
race(5): plus_50 
sex(2): plus_50 capital_loss 
capital_gain(15): plus_50 
capital_loss(15): plus_50 
hours_per_week(6): plus_50 
native_country(42): plus_50 
plus_50(2): 
LogScore Bayes: -454151.06415610015
LogScore BDeu: -458613.8065210351
LogScore MDL: -458373.7686675328
LogScore ENTROPY: -454908.41334505216
LogScore AIC: -455575.41334505216
\end{lstlisting}
